---
title: "Flow analysis exploration"
output: html_notebook
---

```{r include=FALSE}
library(plotly)
library(ggplot2)
library(extrafont)
library(gridExtra)
font_install("fontcm")
loadfonts()
```


## Event logs studied:

```{r echo=FALSE}
logs = read.csv("../logs_summary.csv")
DT::datatable(logs, filter = "top")
```
## Training configuration:

* 2/3 x 1/3 temporal split
* zero bucketing
* aggregate encoding
* XGboost with hyperparam optimization (mainly, # of trees)


## Surveyed techniques:

* predictive flow analysis (FA)
* adaptive FA (based on a variable threshold e.g. 30K samples in the training set )
* mean FA - each cycle time and branching probability is derived based on historical averages
* simple regression - remaining time is predicted as a single value (comparable to [Massimiliano de Leoni's approach](http://www.sciencedirect.com/science/article/pii/S0306437915001313))
* [Wil](https://www.researchgate.net/publication/220503961_Time_prediction_based_on_process_mining) and [Andreas Rogge-Solti](https://link.springer.com/chapter/10.1007/978-3-642-45005-1_27) will be added later as additional baselines. LSTM should probably also be added


```{r include=FALSE}
dat = c()
filenames <- list.files()[grep(paste("^validation_FA_(?=.*\\.csv)", sep=''), list.files(), perl=TRUE)]
for (filename in filenames) {
  this = read.csv(filename, sep=",", header = TRUE)
  this = this[this$metric == "mae",]
  this$nr_cases = this$nr_cases/sum(this$nr_cases)
  this$score_weighted = this$score * this$nr_cases
  this$method2 = "FA"
  this$method3 = strsplit(filename,"_")[[1]][length(strsplit(filename,"_")[[1]])]
  this$filename = filename
  dat = rbind(dat, this)
}
```


```{r include=FALSE}
filenames2 <- list.files()[grep(paste("^validation_direct_(?=.*\\.csv)", sep=''), list.files(), perl=TRUE)]
for (filename in filenames2) {
  this = read.csv(filename, sep=",", header = TRUE)
  this = this[this$metric == "mae",]
  this$nr_cases = this$nr_cases/sum(this$nr_cases)
  this$score_weighted = this$score * this$nr_cases
  this$method2 = "regression"
  this$method3 = strsplit(filename,"_")[[1]][length(strsplit(filename,"_")[[1]])]
  this$filename = filename
  dat = rbind(dat, this)
}

```

```{r echo=FALSE}
size1 = nrow(dat)
dat = dat[complete.cases(dat),]
dat$dataset = droplevels(dat$dataset)
size2 = nrow(dat)
if((size1 - size2)/size2 > 0.01)  {
  print("Warning! More than 1% of incomplete rows")
}
```

```{r include=FALSE}
dat$method3[dat$method3 == "3000000.csv"] = "mean"
dat$method3[dat$method3 == "30.csv" & dat$method2 == "FA"] = "predictive"
dat$method3[dat$method3 == "30.csv" & dat$method2 == "regression"] = "simple"
dat$method3[dat$method3 == "3000.csv"] = "adaptive3000"
dat$method3[dat$method3 == "30000.csv"] = "adaptive30000"

dat$method4 = paste(dat$method3, dat$method2, dat$method, sep = "_")
```


```{r include=FALSE}
dat$score = dat$score / (3600*24) # make days
dat$score_weighted = dat$score_weighted / (3600*24)
```

## Average MAE across all prefix lengths for all datasets and methods:
```{r}
agg_MAE_unw = as.data.frame(aggregate(dat$score,by=list(dataset=dat$dataset, method=dat$method4), mean))
agg_MAE_w = as.data.frame(aggregate(dat$score_weighted,by=list(dataset=dat$dataset, method=dat$method4), sum))
agg_MAE_unw[,3] = round(agg_MAE_unw[,3], 3)
agg_MAE_w[,3] = round(agg_MAE_w[,3], 3)
colnames(agg_MAE_unw)[3] = "Avg_unweighted_MAE_days"
colnames(agg_MAE_w)[3] = "Avg_weighted_MAE_days"
agg_MAE = merge(agg_MAE_w, agg_MAE_unw)
agg_MAE = agg_MAE[with(agg_MAE, order(dataset, Avg_weighted_MAE_days, method = "radix")), ]

DT::datatable(agg_MAE, filter = "top")
```

## Plots MAE vs prefix size
```{r}
for (df in levels(dat$dataset)) {
  p = ggplot(dat[dat$dataset == df,], aes(x = nr_events, y = score, color = method4)) + geom_point() + geom_line() + 
    #facet_wrap(~ dataset, ncol = 3, scales = "free") + 
    ylab("MAE, days") + xlab("Prefix length")+
    ggtitle(df)
    theme(legend.position="top")+
    theme(panel.background = element_rect(fill = 'white', colour = 'black',size=0.5)) + 
    theme(panel.grid.major = element_line(colour = 'lightgrey', size = 0.3))+
  theme(legend.background = element_rect(colour = 'white',size = 0.1, linetype='solid'))+
  theme(legend.title=element_blank())
  print(p)
}
```
## Interactive plots can be accessed [here](https://plot.ly/~verenich/83/)
```{r include=FALSE}
p = ggplot(dat, aes(x = nr_events, y = score, color = method4)) + geom_point() + geom_line() + 
  facet_wrap(~ dataset, ncol = 3, scales = "free") + ylab("MAE, days") + xlab("Prefix length")+
  theme(legend.position="top")+
    theme(panel.background = element_rect(fill = 'white', colour = 'black',size=0.5)) + 
    theme(panel.grid.major = element_line(colour = 'lightgrey', size = 0.3))+
  theme(legend.background = element_rect(colour = 'white',size = 0.1, linetype='solid'))+
  theme(legend.title=element_blank())
api_create(p, "flow-analysis-MAE-500-0660", sharing = "public")
```

```{r include=FALSE}
pdf(file="results_191117-500-0660.pdf",family="CM Roman",width=12,height=10)
print(p)
dev.off()
embed_fonts("results_191117-500-0660.pdf",outfile="results_191117-500-0660.pdf")
```
## To-do:

* [probability calibraion](http://scikit-learn.org/stable/auto_examples/calibration/plot_calibration.html) - may increase prediction accuracy of FA, due to branching probabilities potentially being more reliable.
* repeat experiments for Disco and minit datasets
